% CV ----
% This is just here so I know exactly what I'm looking at in Rstudio when messing with stuff.
\documentclass[11pt,]{article}
\usepackage[sc, osf]{mathpazo}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}




\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{hyperref}
\usepackage{array}

% Now begins the stuff that I added.
% ----------------------------------

% Custom section fonts
\usepackage{sectsty}
\sectionfont{\rmfamily\mdseries\large\bf}
\subsectionfont{\rmfamily\mdseries\normalsize\scshape}


% Make lists without bullets
\renewenvironment{itemize}{
  \begin{list}{}{
    \setlength{\leftmargin}{1.5em}
  }
}{
  \end{list}
}


% Make parskips rather than indent with lists.
\usepackage{parskip}
\usepackage{titlesec}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

% Use fontawesome. Note: you'll need TeXLive 2015. Update.
\usepackage{fontawesome}

% Fancyhdr, as I tend to do with these personal documents.
\usepackage{fancyhdr,lastpage}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0.0pt}
\renewcommand{\footrulewidth}{0.0pt}
\lhead{}
\chead{}
\rhead{}
\lfoot{}
\cfoot{\scriptsize  Eirik Berger Abel - CV}
\rfoot{\scriptsize \thepage/{\hypersetup{linkcolor=black}\pageref{LastPage}}}

% Always load hyperref last.
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref

\hypersetup{unicode=true,
            pdftitle={Eirik Berger Abel:  CV (Curriculum Vitae)},
            pdfauthor={Eirik Berger Abel},
            pdfkeywords={R Markdown, academic CV, template},
            colorlinks=true,
            linkcolor=blue,
            citecolor=blue,
            urlcolor=blue,
            breaklinks=true, bookmarks=true}
\urlstyle{same}  % don't use monospace font for urls

% Make AP style (kinda) dates for the updated/today field

\usepackage{datetime}
\newdateformat{apstylekinda}{%
  \shortmonthname[\THEMONTH]. \THEDAY, \THEYEAR}

% \emph{Updated:} \apstylekinda\today
% ^ removed from that front bar.

\usepackage{orcidlink}
%\usepackage{academicons}
%\definecolor{orcidlogocol}{HTML}{A6CE39}
% \newcommand{\orcid}[1]{\href{https://orcid.org/#1}{\textcolor[HTML]{A6CE39}{\aiOrcid}}}

\begin{document}


\centerline{\huge \bf Eirik Berger Abel}

\vspace{2 mm}

\hrule

\vspace{2 mm}

\moveleft.5\hoffset\centerline{PhD Research Scholar \textbar{} Norwegian
School of Economics}

\moveleft.5\hoffset\centerline{ \faEnvelopeO \hspace{1 mm} \href{mailto:}{\tt \href{mailto:eirik.berger@gmail.com}{\nolinkurl{eirik.berger@gmail.com}}} \hspace{1 mm}  \faPhone \hspace{1 mm}  +47
954 61 029  \hspace{1 mm}         | \emph{Updated:} 07 Jul 2024}



\vspace{2 mm}

\hrule



\hypertarget{education}{%
\section{Education}\label{education}}

PhD in Economics at Norwegian School of Economics (NHH) \hfill 2018 -
ongoing\vspace{-0.2cm}

M.Sc in Business and Administration at Norwegian School of Economics
(NHH) \hfill 2016 - 2018\vspace{-0.2cm}

Bachelor in Business and Administration at Norwegian School of Economics
(NHH) \hfill 2012 - 2015\vspace{-0.2cm}

\vspace{0.5cm}

\hypertarget{visiting-activities}{%
\section{Visiting Activities}\label{visiting-activities}}

Visiting student researcher at Stanford University (visiting Ran
Abramitzky) \hfill 2021 - 2022\vspace{-0.2cm}

\vspace{0.5cm}

\hypertarget{research-fields}{%
\section{Research Fields}\label{research-fields}}

\vspace{-1mm}
\begin{tabular}{p{.85in}>{\hangindent=1em}p{5.65in}<{\raggedright}}
Primary:   & Labor Economics and Applied Microeconomics\\
Secondary: & Economic History\\
\end{tabular}
\vspace{-1mm}

\hypertarget{placement-director}{%
\section{Placement Director}\label{placement-director}}

\begin{tabular}{p{2.15in}>{\hangindent=1em}p{2.15in}>{\hangindent=1em}p{2.15in}<{\raggedright}}
\textbf{Prof. Katrine V. L\o ken}\\
Norwegian School of Economics\\
\href{mailto:Katrine.Loken@nhh.no}{Katrine.Loken@nhh.no}\\
\end{tabular}

\hypertarget{references}{%
\section{References}\label{references}}

\begin{tabular}{p{2.15in}>{\hangindent=1em}p{2.15in}>{\hangindent=1em}p{2.15in}<{\raggedright}}
\textbf{Prof. Kjell Gunnar Salvanes}   & \textbf{Prof. Aline Bütikofer} &  \textbf{Prof. Ran Abramitzky}\\
Norwegian School of Economics & Norwegian School of Economics & Stanford University\\
\href{mailto:Kjell.Salvanes@nhh.no}{Kjell.Salvanes@nhh.no} \vspace{2mm} & \href{mailto:Aline.Buetikofer@nhh.no}{Aline.Buetikofer@nhh.no} \vspace{2mm} & \href{mailto:RanAbr@stanford.edu}{RanAbr@stanford.edu}\\
\end{tabular}

\hypertarget{job-market-paper}{%
\section{Job Market Paper}\label{job-market-paper}}

``\textbf{Living the American Dream: How Norway Became a High-Mobility
Country''}

I estimate long-run trends in intergenerational mobility in income in
Norway for a period that includes World War II and the creation of the
welfare state. I show that persistence between fathers and sons was high
in the early 20th century but decreased substantially for cohorts born
between the 1920s and 1940s. The convergence of incomes between rural
and urban areas explains about half the total fall in persistence.
First, I relate this result to changes in education by using plausibly
exogenous variation in the intensity of schooling from a primary school
reform, which reduced the gap between cities and rural areas, and find
that it significantly decreased persistence in incomes across
generations. Second, I show that the returns to education fell
dramatically at the beginning of World War II. Comparing persistence for
a set of father-son pairs but using income for the father measured just
before and after this shock, I find that the onset of World War II
lowered persistence in income. These results suggest that equal access
to education and a compressed income distribution are two key drivers
behind Norway's transition to high mobility.
\vspace{0.3cm}\vspace{-0.2cm}

\vspace{0.5cm}

\clearpage

\hypertarget{working-papers}{%
\section{Working Papers}\label{working-papers}}

``\textbf{Fertility, Family Formation and Human Capital'' with Aline
Bütikofer and Kjell G. Salvanes \hfill 2024}

This paper generates new insights into the effect of education on
fertility and partner choice across multiple generations. Using an
intensity-of-treatment design, we leverage population-wide panel data
for Norway in combination with a school reform in the 1930s, changing
the instruction time during the school year in rural municipalities. The
reform was binding for most of the rural population and allows us to
estimate the effect of education on fertility behavior across the
life-cycle, partner choice, and spillover effects on the next
generation's fertility. We present robust evidence of reduced total
fertility and an increase in the age at first birth driven by increased
years of education, better labor market outcomes, and mating with
better-educated partners. In addition, the reform also affected the
fertility behavior of the children and decreased fertility rates across
multiple generations. \vspace{0.3cm}\vspace{-0.2cm}

``\textbf{The Making of Equality: How the Second World War Shaped the
Norwegian Income Distributions'' with Ran Abramitzky and Kjell G.
Salvanes \hfill 2024}

Wars and episodes of violent disruption have often been associated with
reductions in economic inequality, but the underlying dynamics and
mechanisms are less understood. This paper explores World War II's
effect on the income distribution in Norway. We use a newly digitized
individual-level panel of tax records and document a large decline in
income inequality in Norway that occurred during and after World War II
and persisted until the 1980s. We descriptively show that the decline is
driven by large increases in the incomes of the bottom 50 percent of the
income distribution and, to a smaller degree, reductions among the very
rich. Using a shift-share design, we show that more than half the
decline in inequality during the war is caused by large increases in the
prices of agricultural products and fish. We further identify the effect
of massive investments in fortifications, airports, and war-related
manufacturing on local labor markets. While the total effects vary by
investment type, the investments disproportionally increased incomes for
workers with low skills or low incomes before the war. Our analysis
provides an alternative narrative by centering on the bottom 50 percent
of income earners, contrasting with studies focusing primarily on the
top income shares and capital destruction during wartime.
\vspace{0.3cm}\vspace{-0.2cm}

``\textbf{Norwegian Tax Returns from the 20th Century: An Automated
Machine Learning Approach'' with Kjell G. Salvanes \hfill 2024}

High-quality administrative tax data has been instrumental in advancing
economic research over the past several decades. However, the scarcity
of data prior to the 1970s has posed a significant obstacle to the study
of the distribution of economic outcomes across people and generations
in historical settings. We introduce a data set extending Norwegian tax
records to the early 1900s, digitizing more than 270,000 pages. The data
set includes most Norwegian regions and data on name, occupation, place
of residence, income, and wealth. We link observations across years to
significantly enhance the usefulness of the data. We argue that the
methodologies employed in this paper offer significant potential for
application to a wider range of sources and countries.
\vspace{0.3cm}\vspace{-0.2cm}

\vspace{0.5cm}

\hypertarget{selected-work-in-progress}{%
\section{Selected Work in Progress}\label{selected-work-in-progress}}

``Becoming Sweden: A 115 Year Journey Towards Wealth and Welfare'' with
Josef Sigurdsson and Arash Nekoei\vspace{-0.2cm}

``Explaining the Rise in Fascism: Evidence from the Nasjonal Samling
Party in Norway'' with Kjell Gunnar Salvanes, Tom Melling and Laura
Khoury\vspace{-0.2cm}

\vspace{0.5cm}

\hypertarget{articles-not-intended-for-publication}{%
\section{Articles Not Intended for
Publication}\label{articles-not-intended-for-publication}}

``Equality before the welfare state: The Norwegian income distribution
1892-1929'', Master Thesis, with Håkon Block Vagle
\hfill 2017\vspace{-0.2cm}

\vspace{0.5cm}

\hypertarget{other-engagements}{%
\section{Other Engagements}\label{other-engagements}}

Journalist at Dagens Næringsliv (financial newspaper) \hfill 2016, 2017
(summer)\vspace{-0.2cm}

Board member at Department of Economics at NHH \hfill 2015, 2017,
2021\vspace{-0.2cm}

Journalist at Finansavisen (financial newspaper) \hfill 2014
(summer)\vspace{-0.2cm}

Board member at Norwegian School of Economics (NHH) \hfill 2016 -
2017\vspace{-0.2cm}

\vspace{0.5cm}

\hypertarget{presentations}{%
\section{Presentations}\label{presentations}}

NHH (internal) \hfill 2023\vspace{-0.2cm}

SOLE, EALE, Stanford University (internal), NHH (internal), Uppsala
University \hfill 2021\vspace{-0.2cm}

NHH (internal), University of Southern Denmark (HEDGE mini-workshop),
the Scandinavian (online) PhD Seminar \hfill 2020\vspace{-0.2cm}

\vspace{0.5cm}

\hypertarget{refereeing}{%
\section{Refereeing}\label{refereeing}}

Journal of Human Resources, Demographic Research\vspace{-0.2cm}

\vspace{0.5cm}

\hypertarget{teaching-and-research-assistance}{%
\section{Teaching and Research
Assistance}\label{teaching-and-research-assistance}}

\hypertarget{teaching-assistant}{%
\subsection{Teaching Assistant}\label{teaching-assistant}}

Inequality, Intergenerational Mobility and Early Childhood. Ph.D level
at UiB. \hfill 2021\vspace{-0.2cm}

Econometrics/Econometric Techniques (\emph{ECN402}). Master level at NHH
\hfill 2020 (2x), 2021\vspace{-0.2cm}

Krakk og Kriser (the History of Business Cycles, \emph{FIE431}). Master
level at NHH \hfill 2019, 2023\vspace{-0.2cm}

Empirical Strategies for Causal Analysis (\emph{ECO433}). Masters level
at NHH \hfill 2019\vspace{-0.2cm}

Økonomi og Psykologi (Behavioural Economics, \emph{VOA40}). Bachelor
level at NHH \hfill 2016\vspace{-0.2cm}

\vspace{0.5cm}

\hypertarget{research-assistant}{%
\subsection{Research Assistant}\label{research-assistant}}

Prof.~Katrine Løken \hfill 2017\vspace{-0.2cm}

Prof.~Alexander Cappelen and Bertil Tungodden \hfill 2016\vspace{-0.2cm}

Prof.~Kjell Gunnar Salvanes \hfill 2017-2023\vspace{-0.2cm}

\vspace{0.5cm}

\hypertarget{skills}{%
\section{Skills}\label{skills}}

\hypertarget{data-science}{%
\subsection{Data science}\label{data-science}}

R (advanced), Stata (advanced), Python (advanced), Bash (proficient) and
Git (proficient)\vspace{-0.2cm}

\vspace{0.5cm}

\vspace{-0.3cm}

\hypertarget{languages}{%
\subsection{Languages}\label{languages}}

Norwegian and English\vspace{-0.2cm}

\vspace{0.5cm}

\end{document}
